\section{Conclusion}
We compared how five disease types spread across a European multilayer transport network, and which
network-science interventions best contain them, holding every factor but the one under study fixed.
Three findings recur across all five types and are resolved against a ten-seed ensemble. First, the
disease type orders the outbreak by dynamical class rather than by any single rate, while climbing
the air $\rightarrow$ $+$water $\rightarrow$ $+$land realism ladder mostly adds people rather than
per-capita severity: the peak fraction stays nearly flat, and ferries even desynchronise it. Second,
immunising the structural bridges (betweenness) beats immunising the busiest hubs (degree) for every
disease, and it is also the most geographically even choice and the one with the widest margin once
mortality rather than peak prevalence is the yardstick. Third, the multilayer substrate changes what
containment means: it makes node targeting \emph{harder}, since land and ferry redundancy routes the
outbreak around any removed city, and it overturns the interdiction conclusion a single-layer model
would reach, where grounding flights ends an air-only epidemic but leaves a third to a half of the
multimodal one standing.

Underlying these results is the network structure itself. The transport substrates are sparse but
strongly small-world, with heavy-tailed degrees and a community organisation that the land and ferry
layers sharpen, and their added bridges surface as low-degree, high-betweenness anomalous gateways:
precisely the cities betweenness targeting protects and degree targeting misses. The study is
comparative rather than predictive, and its absolute counts are not forecasts. The same engine
extends without modification to other regions, finer coverage grids, calibrated parameters, and full
transition-level stochasticity, so the stable core established here, bridges over hubs, realism
raising the cost of containment, and the limits of route closure, is a foundation on which a more
quantitative, region-specific study can build.
